Using cosmic variance to constrain the dark matter halo mass of Lyman-alpha emitting galaxies at $z$=3.1

### 2013-06-20 paper/ms.tex
making quantitative statements

In this section we consider three additional constraints on the models
we accept as succesul. First by taking models with the
highest possible number of mock surveys consistent with
observations. Second by including additional observational constraints
on the occupation fraction for high redshift LAEs. Third by comparing
the angular correlation function against observational constraints by
\cite{Hayashino2004}.

@@

For each model ${\mathcal M}$there are 15 different mock surveys. In the
previous section we presented the models that had at least one (1)
mock survey with $P>0.05$.

Figure \ref{fig:high_success_rate} shows the number of models
that have at least $N_{\rm high-P}$ mocks with $P>0.05$ for the {\texttt
  match} and {\texttt random} methods.  This shows that there are
here are between $80$ to $100$ models with all the 15 realizations with
$P>0.05$. This a reduction of a factor of $\sim 6$ with repect to the
total number of mocks with at least one consistent mock.

Figure \ref{fig:restriction_mock} presents the locii of these models
in the parameter space $M_{\rm min}-M_{\rm max}$ and $M_{\rm
  min}-f_{\rm occ}$ for the {\texttt match} method, the results for
the {\texttt random} method are similar. In this case the models with
$\log_{10}M_{\rm min}< 11.7$ are greatly reduced. This corresponds to
the regions in the parameter space in Figure \ref{fig:landscape} that
already had a low number of consistent mock surveys. From the right
panel in Figure \ref{fig:restriction_mock} one can see that there is
not a strong selection effect on the occupation fraction $f_{\rm
  occ}$.

@@

We impose an additional constraint by requiring that the models
selected in the previous section to be also compatible with the
observational ACF. The measurementens presented in \citep{Yamada2012}
do not present a new measurement of the angular correlation function
(ACF). Instead we use the measurements performed by
\citep{Hayashino2004} on the LAEs observed in the densest field of
SSA22.

It is important to keep in mind that there are some differences
between this work and \citep{Yamada2012}. The color selection by
\cite{Yamada2012} is less stringent compared to the one by
\cite{Hayashino2004}. Also the EW threshold is different,
\cite{Hayashino2004} uses a cut around $154$\AA instead of $190$\AA
used by \citep{Yamada2012}

@@

The procedure is a done through a least square minimizatio is done to
both the mean ACF and the observational ACF reported by
\cite{Hayashino2004} .  The results are shown  in Figure
\ref{figure:correlation_parameters}  in a $\theta_{0}-\gamma$  plane
where the average and standard deviation for each mock  are shown in
comparison with the resut derived from observations.  Error bars in
these figures represents the standard deviation of the ACF over all
the sub-fields. These error bars are larger than the statistical
uncertainty from the fitting procedure on a single field.

We see in the left panel Figure \ref{figure:correlation_parameters} that
the observational ACF measured by \cite{Hayashino2004} (green dot) set aditional 
constrains on the parameters models. Only models with angular-correlation lenght within  
$\theta_{0}\sim[15,23]arcsec$ reproduces observations and this rule out several models.

In order to identify the models that are consistent with observation we
plot if the right panel of figure \ref{figure:correlation_parameters} 
$\theta_{0}$ vs $M_{\rm min}$ and divide the plots into two sets: Those models
with $\log \frac{M_{\rm max}}{M_{\odot}}<12$ (blue dots) and those with 
$\log \frac{M_{\rm max}}{M_{\odot}}>12$ (green dots). The red rectangle includes
the parameter region wich is consistent with the observational constraint in $\theta_{0}$.
With this restriction we find that models with $\log \frac{M_{\rm max}}{M_{\odot}}>12$ and
$\log \frac{M_{\rm min}}{M_{\odot}}>11.1$can be ruled out.

%In general, we observe that the standard
%deviation of the computed ACF in the subfields increases with
%$M_{min}$ following the same trend as in Figure
%\ref{figure:laes_dist}, as a direct consequence of cosmic variance.

In figure \ref{fig:restriction_mock_and_f_occ_corr} we present the prefered
models in the planes $M_{\rm min}-M_{\rm  max}$ and $M_{\rm min}-f_{\rm occ}$
for the {\tt match} method after applying the observational constraints 
in the occupation fraction and the ACF.

Finally, we also present the ACF in the case of the full SSA22
region which has been homogeneously observed by \citep{Yamada2012}. To
this date the observational ACF has not been reported in the
litereature, therefore our calculations can be considered as
predictions.

In Figure X we present the results for the models. The full list of
these correlation functions can be found in the the data repository
for this paper in \verb"github".

%The comparison between the simulated and observed ACFs is also done using
%$\chi^2$ statistic which includes the information on the measurement
%uncertainties and the statistical dispersio from the $15$ mock surveys
%in the simulation

%\begin{equation}
%\chi^{2} = \sum_{\theta_i} \frac{(\xi_{\rm obs}(\theta_i) - \xi_{\rm
%    sim}(\theta_i))^2}{\sigma_{\rm obs}^2(\theta_i) + \sigma_{\rm
%    sim}^2({\theta_i})}, 
%\end{equation}

%where the sum is done over all the $13$ angle values $\theta_i$ where the
%ACF has been computed. In Figure XX we plot this statistic for all the
%models considered. In all cases $\chi^{2}<7$ which is not sufficient
%to consider that the correlation functions from the simulations are
%statistically different from the observational one. The constribution
%of cosmic variance in the simulations together with the statistical
%uncertainties make all tested models compatible with the observational
%constraints.

%From these tests we conclude that the ACF on small fields does not
%provide additional constraints to further select models for halos
%hosting LAEs. The reason is that

%Given that the ACF reported by Hayashino et al en 2004 is taken over
%the densest  field oserved in the SSA22 region by  Yamada et el in
%2012 it is expected that the predicted ACF in the SSA22 region should
%reproduce this observation. In the  left panel of figure
%\ref{figure:correlation_match} we can see the predicted ACFs  an their
%corresponding standard deviation over the seven fields that mock the
%SSA22  region. It can be seen that the model with $M_min=10.6$ seems
%to better reproduce the  Hayashino's ACF and that the corresponding
%field is in fact an overdense field in the SSA22 region covered by
%Yamada et al.

\begin{figure*}
\begin{center}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/power_law_correlation.pdf}
\hspace{5mm}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/mmin_vs_correlation.pdf}
\end{center}
\caption{Left: The fitted parameters  ($\theta_{0}$ vs $\gamma$) of the 
mean ACF  assuming a powerlaw behavior. Each blue dot corresponds to a
match-model while the green dot represent the powerlaw fit of the observational
ACF reported by \citet{Hayashino2004}.  The error bars represents the dispersion
in each parameter over all the mock surveys for a given model and dispersion due
to the observational error for the measured ACF by \citet{Hayashino2004}. Rigth: 
$\theta_{0}$ vs $M_{min}$. The date is subdivided into two blocks $M_{max}<12.0$ 
(blue) and $M_{max}>12.0$ (green). The red rectangle includes the region within
models are consisted with the observational angular-correlation lenght
($15<\theta_{0}<23$) as it can be deduced from the left panel. \label{figure:correlation_parameters}} 
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/mmin_vs_mmax.pdf}
\hspace{5mm}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/mmin_vs_focc.pdf}
\end{center}
\caption{Best models paramter space when both the constraints on the maximal number
  of  consistent mocks and the occupation  fraction $f_{\rm occ}\leq
  0.2$ are included. $M_{\rm min}$ vs $M_{\rm max}$ (left), $M_{\rm min}$ vs $f_{\rm occ}$ (rigth)
  \label{fig:restriction_mock_and_f_occ_corr}} 
\end{figure*} 
%\begin{figure*}
%\begin{center}
%\includegraphics[width=0.46\linewidth,angle=0]{./plots/random_large_correlation_selected_models.pdf}
%\hspace{5mm}
%\includegraphics[width=0.46\linewidth,angle=0]{./plots/random_full_correlation_selected_models.pdf}
%\end{center} 
%\caption{ mean ACFs   and their correponding standard deviation (error
%  bars)  of some selected models in different mass ranges over the 7
%  subfields of the SSA22 field (left) and the entire 12 field sample
%  (rigth) using the  random
%  configuration. \label{figure:correlation_random} }   
%\end{figure*}

@@

%\begin{figure*}
%\begin{center}
%\includegraphics[width=0.49\linewidth,angle=0]{./plots/mytest.pdf}
%\includegraphics[width=0.49\linewidth,angle=0]{./plots/mytest.pdf}
%\end{center} 
%\caption{Spatial distribution for two mocks corresponding to the model
%  $M_{\rm min}=$, $M_{\rm max}=$ and $f_{\rm occ}=$. All the 15
%  different mock  surveys for this model in the {\texttt match}
%  configuration are consistent with observations at the $P>0.05$
%  level. The full data for all the mocks can be found in the github
%  repository for this paper.
%  \label{figure:spatial_distro}.}
%\end{figure*}

Considering the tightest constraints on the mock catalogs 
with the angular correlation function and the maximal
number of consistent mock surveys we end up with 50 models out of the
original 90000 models.

We find that most of the models with an occupation fraction $f_{\rm
  occ}\leq 0.2$ are those with mean correlation lenghts within the
observational uncertainty. Most of the models with $f_{\rm occ}$ are
those that are consistent only when the statistical uncertainty from
the mock cosmic variance is taken into account.

This slight preference towards low occupations fractions is remarkably
close to  the observational constraint $f_{\rm occ}$ derived at
$z=2.2$ by \cite{Hayes2010}.  This estimation was based
on blind surveys of the H$\alpha$ and Lyman $\alpha$ line with the
European Southern  Observatorio (ESO) Very Large Telescope
(VLT). Using corrections by extinction to obtain an estimate for the
intrinsic H$\alpha$ luminosity, and using values for the theoretical
expectation of the ratio Lyman$\alpha$/H$\alpha$ they derive an bulk escape fraction for
the Lyman$\alpha$ radiation of $f_{\rm esc}=(5.3\pm 3.8)\%$ or $f_{\rm
esc}=(10.7\pm 2.8)\%$ if a different dust correction is used. The
authors show that the luminosity function for LAEs at $z=2.2$ is
consistent with the escape fraction being constant for every galaxy
regardless of its luminosity. From this results they derive that
almost $90\%$ of the star forming galaxies emit insufficient
Lyman $\alpha$ to be detected, effectively setting the occupation
fraction to be $f_{\rm occ}=0.10$. For the cosmological parameters
used in this \documentname the age of the universe between $z=3.1$ and
$z=2.2$ has changed by $\sim 1$ Gyr. We assume that the physical
conditions that determine the escape fraction $f_{\rm esc}$ and the
occupation fraction $f_{\rm  occ}$ remain similar over that time
scale.

%that there are 60 set of parameters of our model, out of the original
%90000 initial models, that are consistent with the observational
%constraints at redshift $3.1$: the distribution of the number density,
%the inferred values for the average occupation fraction. The
%consistency with the angular correlation, in terms of the $\chi^2$
%statistics did not help to discard any additional models with a
%significant degree of confidence.

These 50 models can be classified into two families of the same
size. The first, where the range $M_{\rm min}-M_{\rm max}$ is narrow,
typically of less than $<1.0$ dex. While in the second familiy the
extent $>1.0$ dex. In the first case the minimum halo mass is found to
be in a wide range $10^{10}\hMsun <M_{\rm min}< 10^{11.5}\hMsun$ while
in the second case, only models with $M_{\rm min}\sim 10^{10.9}\hMsun$
are compatible with the observational contraints. In what follows we
discuss the implications of the existence of these two families of
models.

@@

\subsection{Observational estimates for the occupation fraction}

We limit the occupation fraction to
be in the range $f_{\rm  occ}\leq 0.2$ to allow for some flexibility on the time evolution
and the uncertainty in the dust correction used to infer the escape fraction. 
Figure  \ref{fig:restriction_mock_and_f_occ}  shows the prefered
models in the planes $M_{\rm min}-M_{\rm  max}$ and $M_{\rm
  min}-f_{\rm occ}$ for the {\tt match} and {\tt   random} methods.
With this additional constraint between $35$ to $40$ models are
consistent with observations, this represents a factor of $\sim 2$
reduction. All the models with $\log_{10}M_{\rm min}\geq 11.0$ are now
excluded and the best models are now clustered around a narrow region
in parameter space. %The list for the model parameters is in the
%Appendix in Table \ref{table:models_match}.

>>>>

In this section we consider two additional constraints on the models
we accept as succesul. First we take models with the
highest possible number of mock surveys consistent with
observations. Second by comparing the angular correlation function
against observational constraints by \cite{Hayashino2004}.

@@

For each model ${\mathcal M}$ there are 15 different mock surveys. In the
previous section we presented the models that had at least one (1)
mock survey with $P>0.05$.

Figure \ref{fig:high_success_rate} shows the number of models
that have at least $N_{\rm high-P}$ mocks with $P>0.05$ for both the
{\texttt  match} and {\texttt random} methods.  This shows that there
are between $80$ to $100$ models with all the 15 realizations with
$P>0.05$. Selction of these models as successful represents a
reduction of a factor of $\sim 6$ with repect to the total number of
mocks with at least one consistent mock.

Figure \ref{fig:restriction_mock} presents the locii of these models
in the parameter space $M_{\rm min}-M_{\rm max}$ and $M_{\rm
  min}-f_{\rm occ}$. The results are very similar between the {\texttt
  match} and {\texttt random} methods. With this constraints the
number of consistent models with  $\log_{10}M_{\rm min}< 11.7$ are
greatly reduced. This corresponds to the regions in the parameter
space in Figure \ref{fig:landscape} that already had a low number of
consistent mock surveys. On the other hand, from the right panel in
Figure \ref{fig:restriction_mock} one can see that there is not a
strong selection effect on the occupation fraction $f_{\rm occ}$.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/power_law_correlation.pdf} 
\hspace{5mm}  
\includegraphics[width=0.46\linewidth,angle=0]{./plots/mmin_vs_correlation.pdf} 
\end{center}
\caption{Left: Values for the free parameters ($\theta_{0}$ vs $\gamma$) of the 
mean angular correlation function of the 15 mock surveys for each
model. Blue dots corresponds to simulations and the green cross to
observations by \citet{Hayashino2004}. The error bars in the
theoretical data correspond to the starndard deviation from the
different mocks. Right: Same results as before in the
$\theta_{0}$-$M_{\rm min}$ plane. This time the observational
constraints are represented by the red rectangle. Blue squares
represent models with $\log_{10} M_{\rm  max}<12.0$, green circles are
models with $\log_{10} M_{\rm max}>12.0$.
\label{figure:correlation_parameters}}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/mmin_vs_mmax.pdf}
\hspace{5mm}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/mmin_vs_focc.pdf}
\end{center}
\caption{Best models paramter space when both the constraints on the
  maximal number of  consistent mocks and the occupation  fraction
  $f_{\rm occ}\leq  0.2$ are included. $M_{\rm min}$ vs $M_{\rm max}$
  (left), $M_{\rm min}$ vs $f_{\rm occ}$ (rigth).
  \label{fig:restriction_mock_and_f_occ_corr}} 
\end{figure*}

As an additional observational constraint we requiring that the models
selected in the previous section must be compatible with the
observational angular correlation function (ACF). The measurementens
presented in \citep{Yamada2012} do not report an ACF
measurement. Instead we use the results reported by
\citep{Hayashino2004} on the LAEs observed in the densest field of
SSA22.

It is important to keep in mind that there are some differences
between this work and \citep{Yamada2012}. First, the color selection by
\cite{Yamada2012} is less stringent compared to the one by
\cite{Hayashino2004}. Also the EW threshold is different,
\cite{Hayashino2004} uses a cut around $154$\AA instead of $190$\AA
used by \citep{Yamada2012}

@@

Both the theoretical and observational ACF are derived by a least
square minimization procedure. The results are shown  in Figure
\ref{figure:correlation_parameters}  in a $\theta_{0}-\gamma$  plane
where the average and standard deviation for each mock is shown in
comparison with the resut derived from observations.  The Error bars in
these figures represents the standard deviation of the ACF over all
the sub-fields in the 15 mock observations. These error bars are
larger than the statistical uncertainty from the fitting procedure on
a single field..

We see in the left panel Figure \ref{figure:correlation_parameters} that
the observational ACF measured by \cite{Hayashino2004} (green dot) is
succesul in reducing the total number of possible models. Only those
with angular-correlation length within
$15^{\prime\prime}<\theta_{0}<23^{\prime\prime}$ are consideret to
reproduce observations.

In order to interpret the results from this test we
show in the right panel of Figure \ref{figure:correlation_parameters} 
the plane $\theta_{0}$-$M_{\rm min}$ for all models. There we divide
the models into two disjoint sents: those with $\log \frac{M_{\rm
    max}}{M_{\odot}}<12$ (blue dots) and $\log \frac{M_{\rm
    max}}{M_{\odot}}>12$ (green dots). The colored rectangle includes
the parameter region wich is consistent with the observational
constraint in $\theta_{0}$. With this restriction we find that most
models with $\log \frac{M_{\rm max}}{M_{\odot}}>12$ and  $\log
\frac{M_{\rm     min}}{M_{\odot}}>11.1$can be safely ruled out.

In Figure \ref{fig:restriction_mock_and_f_occ_corr} we present the prefered
models in the planes $M_{\rm min}-M_{\rm  max}$ and $M_{\rm min}-f_{\rm occ}$
for the {\tt match} method after applying the observational constraints 
in the occupation fraction and the ACF. Here we consider a model
consistent if there is an overlap of $1-sigma$ in its values for the
correlation length $\theta_0$ and the power $\gamma$ in the
correlation function.

@@

Imposing additional constraints with the maximal number of consistent
mock surveys and the angular correlation function we end up with 50
models out of the original 90000 models. In what follows we discuss
the implications of the selection of these models on the escape
fraction of Lyman $\alpha$ radiation, the stellar abundance and the
star formation efficiency in LAEs at $z=3.1$.

\subsection{Escape fraction}

We find that the majority ($80\%$) of the best models have a low
ocacupation fraction $f_{\rm occ}\leq 0.4$. This preference goes in
the same direction as the observational constraint on $f_{\rm occ}\sim
0.1-0.2$ derived at $z=2.2$ by \cite{Hayes2010}.

Their estimation was based on blind surveys of the H$\alpha$ and Lyman $\alpha$ line with the
European Southern  Observatorio (ESO) Very Large Telescope
(VLT). Using corrections by extinction to obtain an estimate for the
intrinsic H$\alpha$ luminosity, and using values for the theoretical
expectation of the ratio Lyman$\alpha$/H$\alpha$ they derive an bulk escape fraction for
the Lyman$\alpha$ radiation of $f_{\rm esc}=(5.3\pm 3.8)\%$ or $f_{\rm
esc}=(10.7\pm 2.8)\%$ if a different dust correction is used. The
authors show that the luminosity function for LAEs at $z=2.2$ is
consistent with the escape fraction being constant for every galaxy
regardless of its luminosity. From this results they derive that
almost $90\%$ of the star forming galaxies emit insufficient
Lyman $\alpha$ to be detected, effectively setting the occupation
fraction to be $f_{\rm occ}=0.10$. For the cosmological parameters
used in this \documentname the age of the universe between $z=3.1$ and
$z=2.2$ has changed by $\sim 1$ Gyr. We assume that the physical
conditions that determine the escape fraction $f_{\rm esc}$ and the
occupation fraction $f_{\rm  occ}$ remain similar over that time
scale.

Including the information of the halo mass range size allows us

These 50 models can be classified into two families of the same
size.

The first, where the range $M_{\rm min}-M_{\rm max}$ is narrow,
typically of less than $<1.0$ dex. While in the second familiy the
extent $>1.0$ dex. In the first case the minimum halo mass is found to
be in a wide range $10^{10}\hMsun <M_{\rm min}< 10^{11.5}\hMsun$ while
in the second case, only models with $M_{\rm min}\sim 10^{10.9}\hMsun$
are compatible with the observational contraints. In what follows we
discuss the implications of the existence of these two families of
models.

@@
